Diagnosising the sensitivity of grounding line flux to changes in sub-ice shelf melting

### 2019-04-22 draft_v1.tex
Updates from Overleaf

\abstract{Motivated by previous work using ice flow models to quantify ice shelf buttressing and its impacts on the flux of ice across the grounding line \citep[e.g.,][]{furst2016,reese2018}, we seek a better physical understanding for how ice dynamics link small ice thickness perturbations, via changes in sub-ice shelf melting, to changes in ice shelf buttressing and ice flux across the grounding line. More specifically, we seek to define one or more ice shelf buttressing ``metrics'' that are readily calculated from standard ice sheet model outputs and are simultaneously informative for diagnosing the sensitivity of grounding line flux to ice thickness at specific locations on an ice shelf. By studying the ice dynamics for both idealized (MISMIP+) and realistic (Larsen C) ice shelves, we find that the first principle stress at perturbation locations is the best overall metric for linking local changes in ice shelf dynamics with changes in the integrated grounding line flux. Unfortunately, this metric only shows a robust relationship with the integrated grounding line flux for regions near the center of an ice shelf; for points too near the grounding line or too near the calving front, no clear relationship exists between any of the readily calculable metrics explored here and changes in grounding line flux. This motivates our exploration of an adjoint-based method for defining grounding line flux sensitivity to local changes in ice shelf geometry. Using the same idealized and realistic test cases, we demonstrate that this method is equivalent to the sensitivity analysis of \citep{reese2018} but requires only a single model adjoint solve. Thus we suggest that the adjoint-based method can provide a model run-time means of analyzing grounding line flux sensitivity to changes in sub-ice shelf melting.}

@@

Marine ice sheets like West Antarctica (and to a lesser extent, portions of East Antarctica) are grounded below sea level and their bedrock would remain so even after full isostatic rebound \citep{barletta2018}. This and the fact that ice sheets generally thicken inland lead to a geometric configuration that is unstable; a small increase in flux at the grounding line thins the ice there, leading to floatation, a retreat of the grounding line into deeper water, further increases in flux (due to thicker ice), and further thinning and grounding line retreat. This theoretical ``marine ice sheet instability'' mechanism \citep{mercer1978, schoof2007} is supported by idealized \citep{schoof2012, asay2016} and realistic ice sheet modeling \citep{royston2016} experiments and some studies \citep{joughin2014,rignot2014} argue that such an instability is currently under way along outlet glaciers of Antarctica's Amundsen Sea Embayment (ASE). The relevant perturbation for grounding line retreat in the ASE is thought to be intrusions of relatively warm, intermediate depth ocean waters onto the continental shelves [ref to recent review papers in Oceanography? or other recent reviews? {\bf{TZ: perhaps Xylar can help with this?}}], which have reduced the thickness and extent of marginal ice shelves via increased submarine melting [REFS]. These reductions are critical because fringing ice shelves restrict the flux of ice across their grounding lines farther upstream -- the so-called ``buttressing'' affect of ice shelves \citep{gudmundsson2012, gudmundsson2013, derydt2015} -- which makes them a critical control on ice flux from Antarctica to the ocean.

@@

%Warm ocean water is getting more evident of being responsible of increased melt at the base of ice shelves in Antarctica \citep{jenkins2018,rintoul2016}. The basal melting decreases ice thickness and reduce the buttressing ability of ice shelf \citep{gagliardini2010}, and can therefore potentially destabilize marine ice sheets. Satellite altimetry measurements have proved this mechanism that ice shelf thickness decrease links glacier acceleration in Antarctica \citep{pritchard2012}. Recent studies also suggested that the thinning of ice shelfves also contributes to the ice dynamic changes of upstream grounded ice sheets \citep{minchew2018}, indicating the importance of ice shelf-ocean coupling in the whole ice sheet dynamics. 
On ice shelves, gradients in hydrostatic pressure are balanced by the primarily extensional flow of ice towards the calving front \citep{hutter1983, morland1987, schoof2007} and, in theory, a one horizontal dimension ($x$-$z$) ice shelf provides no buttressing \citep{schoof2007,gudmundsson2013}. (\textit{SP: I wonder if this requires a monotonic decrease in ice shelf thickness though? \bf{TZ: I think so}}) For realistic, three-dimensional ice shelves however,  buttressing results from three main sources: 1) compressive ice flow 2) lateral shear, and 3) ``hoop'' stress \citep{pegler2012}. Both compressive and lateral shear stresses supply backward resistences to extentional ice shelf flow, and the ``hoop'' stress is a transverse stress arising from azimuthal extension existing in the diverging ice flow regions \citep{wearing2016}. (\textit{For completeness, we should probably briefly describe how each of these contributes to buttressing rather than assume people already know (?). The first two are easy. I'm not sure about the 3rd. \bf{TZ: I add a sentence in the previous line}}) Due to the complex geometries, kinematics, and dynamics of real ice shelves, an understanding of the specific processes and locations that control ice shelf buttressing is far from straigtforward.

Several recent studies apply whole-Antarctic ice sheet models optimized to present-day observations towards improving our understanding for how Antarctica ice shelves limit flux across the grounding line and, by extension, ice dynamics farther inland. \cite{furst2016} calculated the buttressing across Antarctica ice shelves along two major directions (flow and second principle stress) and evaluated their impacts on upstream ice dynamics to identify regions of the ice shelves that are dynamically ``passive'', such that increased submarine melting, or even complete removal of ice in these areas should not significantly alter local or regional ice dynamics or the flux of ice upstream. \cite{reese2018} used perturbation experiments to link small, localized decreases in ice shelf thickness to changes in integrated grounding line flux (GLF), thereby providing a map of GLF sensitivity to local increases in submarine melt rates. 
%SP: After writing this sentence, it occurs to me that something else that we could follow up on at some point is if / how broader patterns of ice thickness change might complicate the interpretation of these simple perturbation experiments. E.g., you can imagine that 1) multiple perturbations in diff. areas of the ice shelf might conflict with one another and have a more complex impact on g.l. flux (e.g., simultaneous perturbations in multiple areas that show both g.l. flux increases and decreases w/ ice thickness perturbations), 2) similar scale perturbations but applied over wider areas (e.g., over an area that is representative of the actual length scale of submarine melt patterns). Maybe this would be implicit in any 2nd or later study where we perturbed ice shelves based on the sensitivity maps we derive from the adjoint method.

@@

Below, we first provide a brief description of the ice sheet model used in our study. We follow with a description of the model experiments and a discussion of the experimental results and their interpretation. We then demonstrate and discuss the pros and cons of a number of possible metrics for quantifying GLF sensitivity to changes in submarine melt. Based on limitations in all metrics explored here, we conclude by proposing and demonstrating an adjoint-based calculation that provides a sensitivity map analogous to that from the \cite{reese2018} perturbation experiments but at the cost of a single model adjoint solve. ({\bf{TZ: we should revisit this paragraph at some point after we finish most of revisions in the manuscript}})

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig1.pdf}
\caption{Plan view of steady-state surface ice speeds for MISMIP+ (a) and present-day surface ice speed for Larsen C ice shelf (b). The white curves show the grounding lines.}
\label{fig1}
\end{figure}

@@

Here, we apply MALI to experiments on both idealized and realistic marine-ice sheet geometries. For our idealized domain and model state, we start from the equilibrium initial conditions for the MISMIP+ experiments, as described in \cite{asay2016} and Cornford and others (MISMIP+ papers) ({\bf{TZ: is it the same paper as Xylar's?}}). The model mesh is spatially uniform at 2 km resolution. For our realistic domain, we use Antarctica's Larsen C ice shelf and its upstream catchement area. The model state is based on the optimization of the ice stiffness ($\gamma$ in Equation \ref{eq:effvisc}) and basal friction ($\beta$ in Equation \ref{eq:basalbc}) coefficients to match present-day velocities \citep{rignot2014} using adjoint-based methods discussed in \cite{perego2014} and \cite{hoffman2018}. The domain geometry is based on BEDMAP2 \citep{fretwell2013} and ice temperatures, which are held fixed for this study, are based on \cite{liefferinge2013}. Mesh resolution coarsens to 20 km in the ice sheet interior and is no greater than 6 km within the ice shelf \textit{This seems coarse to me ... don't we go to finer resolution, e.g. 2 km, near the g.l.?} {\bf{TZ: I think it means the other way around -- the res. is not coarser than 6 km in ice shelf (i.e., $>$2 km and $<$ 6 km)}}. Following optimization to present-day velocities, the model is relaxed using a 100 year forward run and it is this initial condition from which the experiments discussed below are conducted. Both the MISMIP+ and Larsen C experiments use 10 vertical layers that are finest near the bed and coarsen towards the surface (4\% and 23\% of the total thickness, respectively). The grounding line position is determined from hydrostatic equilibrium and a sub-element parameterization analogous to method SEP3 from \cite{seroussi2014} is used to define basal friction coefficient values at the grounding line.

@@

To explore the sensitivity of changes in GLF to small changes in ice shelf thickness, we conduct a number of perturbation experiments analagous to those of \citet{reese2018}. Using diagnostic model solutions, we first study the instantaneous response of GLF for the idealized geometry and initial state provided by the MISMIP+ experiment \citep{asay2016}. We then conduct a similar study but using a realistic configuration and initial state for Antarctica's Larsen C ice shelf. %{\bf{possibly add more real ice shelves?}})

Our experiments are conducted in a manner similar to those of \cite{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, the uniform, 2 km mesh implies that grid cell centers naturally align with these boxes. For the Larsen C ice shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we use 2$\times$2 km square boxes that are also the real cell size. For Larsen C, we only use 20$\times$20 km square boxes (i.e., as in \citet{reese2018}). Lastly, for the MISMIP+ 2 km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is likely lateraly buttressed) and $y>40$ km (one half of the ice shelf due to the symmetry about the centerline).

Similar to \cite{reese2018}, we define a GLF response number
\begin{equation}
N_r = \left(\frac{R}{P}\right)^k,
\end{equation}
where $R$ is the ice flux change integrated along the entire grounding line, $P$ is the mass associated with a single grid box perturbation (e.g.,2 km$\times$2 km $\times$0.001 km for the MISMIP+ perturbation experiments) and $k$ is a power-law index that allows for the possiblity of a nonlinear relationship between ice shelf buttressing and the change in GLF (see also \cite{schoof2007}). Here, we use $k=1/n$ ($n=3$).

Despite the existence of many different factors (ice flow directions, horizontal gradients of ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc), we here mainly present the results of the stress fields and the distances between perturbations and GL, as they appears to correlate more closely to the sensitivity of GL flux change to ice shelf perturbations. Similar to  \cite{furst2016} (Eqn \ref{butN}), we calculate buttressing numbers ($N_b$) as follows,
\begin{equation}
N_b=1-\frac{\sigma_{nn}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=0.5\left(1-{\rho_i}/{\rho_w}\right)gH$). $\rho_i$ (910 kg m$^3$) and $\rho_w$ (1028 kg m$^3$) are the density of ice and ocean water, respectively. $\sigma_{nn}$ is the normal stress along certain horizontal direction.

@@

\section{Results and Discussions}
\subsection{Linear relationship between buttressing ($N_b$) and GL flux responses ($N_r$)}
The decrease of ice shelf buttressing tends to induce the increase of GLF \citep{gagliardini2010}. Therefore, the highly buttressed regions are in general more sensitive to sub-ice shelf melting for  marine ice sheet dynamics. It is thus useful to understand more quantitatively the relationship between the buttressing ``strength'' (number) and the GLF  changes. Can we predict the changes of GLF simply by the buttressing number ($N_b$)? By calculating the buttressing number ($N_b$) along the first principle stress ($\sigma_{p1}$) direction, we can see very weak $N_b$-$N_r$ correlations from the results of all perturbation experiments (Fig \ref{fig2}b). However, if we do not consider the perturbation points that are weakly buttressed ($x>480$ km; near the open shelf region) and are close to GL (the minimum distances between GL and perturbation points are greater than 5 km), we can clearly see a strong linear $N_b$-$N_r$ regression (Fig \ref{fig2}c). A similar strong linearity can also be found if we apply a different filtering approach using the ice thickness gradient field (Fig \ref{fig2}a), i.e., we keep the perturbation points that have thickness gradients smaller than 0.0007 (Fig \ref{fig2}d). This nonlinearity feature is likely caused by the intensive shearing and complex ice flow mechanics near GL (a strong transition zone from relatively slow grounded ice flow to fast extensively floating ice flow). We can again see this non-linearity region in our adjoint sensitivity experiments (Fig. \ref{fig9} in section ``Adjoint sensitivity'').

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig2.pdf}
    \caption{(a) The magnitude of ice thickness gradient around GL (white curve) for the steady-state MISMIP+ geometry; (b) The relationship of $N_b$-$N_r$ for all perturbation points; (c) The relationship of $N_b$-$N_r$ for perturbation points that are $>$10 km away from GL; (d) The relationship of $N_b$-$N_r$ for perturbation points with thickness gradient $<$7$times$10$^{-3}$. The red (blue) dots in the insets are GL (perturbation) cells.}
\label{fig2}
\end{figure}

\subsection{The buttressing directions}
According to Equation \ref{butN}, the buttressing number $N_b$ is computed by the normal stress ($\sigma_{nn}$) along specific directions. Therefore, the buttressing at certain perturbation points can be various for different directions we apply. In \cite{furst2016} the buttressing along two significant directions (flow and second principle stress ($\sigma_{p2}$)) was analyzed, and found that the $\sigma_{p2}$ buttressing, which points to the most compression direction, has the maximum impacts on the ``passive'' ice shelf regions. Here we get the correlation numbers ($r^2$) between $N_r$ and $N_b$ for different directions with respect to the first principle stress and flow direction, respectively (Fig \ref{fig3}). Differently, we find that $N_b$ along the $\sigma_{p1}$ direction shows the best regression performance, whereas the $\sigma_{p2}$ appears to show the weakest correlations to $N_r$. This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$). From Figure \ref{fig3}b, we see that for around 50\% of the perturbation points, their flow directions are around 30--50 degree more than the $\sigma_{p1}$ directions, which is consistent with the phase differences in Figure \ref{fig3}a. If we add around 40 degree on top of the flow direction (blue curve), the new direction will be likely aligned closely with the $\sigma_{p2}$ direction, which is consequently pointing to the smallest $r^2$ number for the $\sigma_{p1}$ (red) curve. This indicates that the local maximum buttressing relating to $\sigma_{p2}$ is unnecessarily corresponding to the integrated instantaneous GLF responses.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig3.pdf}
    \caption{(a) The $N_b$-$N_r$ regression coefficients for each direction rotated anti-clock-wisely from the $\sigma_{p1}$ (red) and flow direction (blue); (b) The histogram of the angle differences between flow and $\sigma_{p1}$ directions. The perturbation points we apply here are the same as in Figure \ref{fig2}d.}
    \label{fig3}
\end{figure}

\subsection{Possible controlling factors}
Further clues of the impacts from different chosen directions can be found in Figure \ref{fig4}. We analyzed the number of maximum (light red) and minimum (blue) velocity changes around each perturbation point for the cases using flow (Fig \ref{fig4}a) and $\sigma_{p1}$ directions (Fig \ref{fig4}b). Figure \ref{fig4}a and c contain all perturbation points while Figures \ref{fig4}b and d only include the filtered perturbation points as in Figure \ref{fig2}d. For the flow direction results, most of the maximum (minimum) velocity change events occur along the flow (120 degree) direction. The $\sigma_{p1}$ direction results are more spread than of the flow direction results. However, it is still clear that most of the maximum (minimum) velocity change events are aligned near the first (second) principle stress direction, a supporting evidence for our previous findings.

\begin{figure}
	\centering
    \includegraphics[width=1\linewidth]{figs/fig4.pdf}
    \caption{Histograms of the frequency of the neighboring maximum (red) and minimum (blue) velocity change around all perturbation points (a and c) and around the same selected perturbation points (b and d) as Figure \ref{fig2}d ({\bf{TZ: this figure still needs double check!}})}
	\label{fig4}
\end{figure}

Another evidence is the standard deviation of the differences between the relative changes of stress and velocity ($\Psi$) along GL (Fig \ref{fig5}). 
\begin{equation}
\Psi = \frac{\sigma_{p}-\sigma_{c}}{\sigma_{c}} - \frac{u_{p}-u_{c}}{u_{c}},
\label{psi}
\end{equation}
where the subscripts $p$ and $c$ denote perturbation experiments and ctrl runs. $\sigma$ and $u$ denote the stress component and ice velocity, respectively. This is a measure of the consistency of velocity and stress variances along GL under certain perturbation experiments. Here we choose $\sigma_{p1}$ and $\sigma_{p2}$ for instances, i.e., we set $\sigma$ in Equation \ref{psi} to $\sigma_{p1}$ and $\sigma_{p2}$. Overall, the $\sigma_{p1}$ case (Fig \ref{fig5}a) shows smaller standard deviations, especially for perturbation points close to GL, than the $\sigma_{p2}$ case (Fig \ref{fig5}b). This possibly indicates that GLF are more relevant to the changes of $\sigma_{p1}$ along GL, compared to the $\sigma_{p2}$ case, which can be clearly found if we look at some specific perturbation example as shown in Figure \ref{fig6}.

>>>>

\abstract{Motivated by previous work using ice flow models to quantify ice shelf buttressing and its impacts on the flux of ice across the grounding line \citep[e.g.,][]{furst2016,reese2018}, we seek an improved physical understanding for how ice dynamics link small ice thickness perturbations, via changes in sub-ice shelf melting, to changes in ice shelf buttressing and ice flux across the grounding line. More specifically, we seek to define one or more ``metrics'' that are 1) readily calculated from standard ice sheet model outputs and 2) informative for diagnosing the sensitivity of grounding line flux to changes in ice thickness at specific locations on an ice shelf. By studying the ice dynamics for both idealized (MISMIP+) and realistic (Larsen C) ice shelves, we find that the first principle stress is the best overall metric for linking local changes in ice shelf thickness and dynamics with changes in the integrated grounding line flux. Unfortunately, this metric only shows a robust relationship with the integrated grounding line flux for regions near the center of an ice shelf; for points too near the grounding line or the calving front, no clear relationship exists between any of the readily calculable metrics explored here and changes in grounding line flux. This motivates our exploration of an adjoint-based method for defining grounding line flux sensitivity to local changes in ice shelf geometry. Using the same idealized and realistic test cases, we demonstrate that this method is equivalent to the sensitivity analysis of \citep{reese2018} but requires only a single model adjoint solve. We conclude that the adjoint-based method can provide a means of analyzing grounding line flux sensitivity to changes in sub-ice shelf melting at model run time.}

@@

Marine ice sheets like West Antarctica (and to a lesser extent, portions of East Antarctica) are grounded below sea level and their bedrock would remain so even after full isostatic rebound \citep{barletta2018}. This and the fact that ice sheets generally thicken inland lead to a geometric configuration that is unstable; a small increase in flux at the grounding line thins the ice there, leading to floatation, a retreat of the grounding line into deeper water, further increases in flux (due to thicker ice), and further thinning and grounding line retreat. This theoretical ``marine ice sheet instability'' mechanism \citep{mercer1978, schoof2007} is supported by idealized \citep{schoof2012, asay2016} and realistic ice sheet modeling \citep{royston2016} experiments and some studies \citep{joughin2014,rignot2014} argue that such an instability is currently under way along outlet glaciers of Antarctica's Amundsen Sea Embayment (ASE). The relevant perturbation for grounding line retreat in the ASE is thought to be intrusions of relatively warm, intermediate depth ocean waters onto the continental shelves, which have reduced the thickness and extent of marginal ice shelves via increased submarine melting \citep[e.g.,][]{JenkinsEtAl2016}. These reductions are critical because fringing ice shelves restrict the flux of ice across their grounding lines farther upstream -- the so-called ``buttressing'' affect of ice shelves \citep{gudmundsson2012, gudmundsson2013, derydt2015} -- which makes them a critical control on ice flux from Antarctica to the ocean.

@@

%Warm ocean water is getting more evident of being responsible of increased melt at the base of ice shelves in Antarctica \citep{jenkins2018,rintoul2016}. The basal melting decreases ice thickness and reduce the buttressing ability of ice shelf \citep{gagliardini2010}, and can therefore potentially destabilize marine ice sheets. Satellite altimetry measurements have proved this mechanism that ice shelf thickness decrease links glacier acceleration in Antarctica \citep{pritchard2012}. Recent studies also suggested that the thinning of ice shelfves also contributes to the ice dynamic changes of upstream grounded ice sheets \citep{minchew2018}, indicating the importance of ice shelf-ocean coupling in the whole ice sheet dynamics. 
On ice shelves, gradients in hydrostatic pressure are balanced by the primarily extensional flow of ice towards the calving front \citep{hutter1983, morland1987, schoof2007} and, in theory, a one  dimensional ($x$-$z$) ice shelf provides no buttressing \citep{schoof2007,gudmundsson2013}\footnote{The ice shelf is also assumed to be monotonically decreasing in thickness from the grounding line to the calving front.}. For realistic, three-dimensional ice shelves however,  buttressing results from three main sources: 1) compressive ice flow 2) lateral shear, and 3) ``hoop'' stress \citep{pegler2012}. Both compressive and lateral shear stresses can supply backward resistance to extentional ice shelf flow, and the ``hoop'' stress is a transverse stress arising from azimuthal extension in regions of diverging flow \citep{wearing2016}.  Due to the complex geometries, kinematics, and dynamics of real ice shelves, an understanding of the specific processes and locations that control ice shelf buttressing is far from straightforward.

Several recent studies apply whole-Antarctic ice sheet models optimized to present-day observations towards improving our understanding for how Antarctic ice shelves limit flux across the grounding line (and by extension how they impact ice dynamics farther inland). \cite{furst2016} calculated the buttressing across Antarctica ice shelves along two major directions (aligned with the ice flow and the second principle stress) and evaluated their impacts on upstream ice dynamics to identify regions of the ice shelves that are dynamically ``passive''; in these regions increased submarine melting, or even complete removal of ice in these areas should not significantly alter local or regional ice dynamics or the flux of ice upstream. \cite{reese2018} used perturbation experiments to link small, localized decreases in ice shelf thickness to changes in integrated grounding line flux (GLF), thereby providing a map of GLF sensitivity to local increases in submarine melt rates. \textbf{Add some discussion here about the recent Goldberg et al. paper as well?}

%SP: After writing this sentence, it occurs to me that something else that we could follow up on at some point is if / how broader patterns of ice thickness change might complicate the interpretation of these simple perturbation experiments. E.g., you can imagine that 1) multiple perturbations in diff. areas of the ice shelf might conflict with one another and have a more complex impact on g.l. flux (e.g., simultaneous perturbations in multiple areas that show both g.l. flux increases and decreases w/ ice thickness perturbations), 2) similar scale perturbations but applied over wider areas (e.g., over an area that is representative of the actual length scale of submarine melt patterns). Maybe this would be implicit in any 2nd or later study where we perturbed ice shelves based on the sensitivity maps we derive from the adjoint method.

@@

Below, we first provide a brief description of the ice sheet model used in our study. We follow with a description of the model experiments and a discussion of the experimental results and their interpretation. We then demonstrate and discuss the pros and cons of a number of possible metrics for quantifying GLF sensitivity to changes in submarine melt. Based on limitations in all metrics explored here, we conclude by proposing and demonstrating an adjoint-based calculation that provides a sensitivity map analogous to that from the \cite{reese2018} perturbation experiments but at the cost of a single model adjoint solve. ({\bf{TZ: we should revisit this paragraph at some point after we finish most of revisions in the manuscript. SP: Agreed. It could be that some of the correlations you show w/ just the first principle stress are also worthy of highlighting as being reasonably good too.}})

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig1.pdf}
\caption{Plan view of steady-state surface speed for MISMIP+ (a) and Larsen C ice shelf (b). The white curves show the grounding lines.}
\label{fig1}
\end{figure}

@@

Here, we apply MALI to experiments on both idealized and realistic marine-ice sheet geometries. For our idealized domain and model state, we start from the equilibrium initial conditions for the MISMIP+ experiments, as described in \cite{asay2016} and Cornford and others (MISMIP+ papers) ({\bf{TZ: is it the same paper as Xylar's? SP: No, I meant the actual MISMIP+ results paper, which Steph C. is supposed to be writing.}}). The model mesh is spatially uniform at 2 km resolution. For our realistic domain, we use Antarctica's Larsen C ice shelf and its upstream catchement area. The model state is based on the optimization of the ice stiffness ($\gamma$ in Equation \ref{eq:effvisc}) and basal friction ($\beta$ in Equation \ref{eq:basalbc}) coefficients in order to provide a best match between modeled and observed present-day velocities \citep{rignot2014} using adjoint-based methods discussed in \cite{perego2014} and \cite{hoffman2018}. The domain geometry is based on BEDMAP2 \citep{fretwell2013} and ice temperatures, which are held fixed for this study, are based on \cite{liefferinge2013}. Mesh resolution on the ice shelf is between 2 and 6 km and coarsens to 20 km in the ice sheet interior. Following optimization to present-day velocities, the model is relaxed using a 100 year forward run, providing the initial condition from which the Larsen C experiments (discussed below) are conducted. Both the MISMIP+ and Larsen C experiments use 10 vertical layers that are finest near the bed and coarsen towards the surface (4\% and 23\% of the total thickness, respectively). The grounding line position is determined from hydrostatic equilibrium. A sub-element parameterization, analogous to method SEP3 from \cite{seroussi2014}, is used to define basal friction coefficient values at the grounding line.

@@

To explore the sensitivity of changes in GLF to small changes in ice shelf thickness, we conduct a number of perturbation experiments analagous to those of \citet{reese2018}. Using diagnostic model solutions, we first study the instantaneous response of GLF for the idealized geometry and initial state provided by the MISMIP+ experiment \citep{asay2016}. We then conduct a similar study for Antarctica's Larsen C ice shelf but using a realistic configuration and initial state. %{\bf{possibly add more real ice shelves?}})

Our experiments are conducted in a manner similar to those of \cite{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, the uniform, 2 km mesh implies that grid cell centers naturally align with these boxes. For the Larsen C ice shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we use 2$\times$2 km square boxes that coincide with the actual grid cell size. For Larsen C, we only use 20$\times$20 km square boxes (i.e., as in \citet{reese2018}). Lastly, for the MISMIP+ 2 km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is likely lateraly buttressed) and $y>40$ km (one half of the ice shelf due to the symmetry about the centerline). While we only perturb the ice shelf over one half of the centerline, we analyze the response to those perturbations over the entire model domain.

Similar to \cite{reese2018}, we define a GLF ``response number'',
\begin{equation}
N_r = \left(\frac{R}{P}\right)^k,
\end{equation}
where $R$ is the ice flux change integrated along the entire grounding line, $P$ is the mass associated with a single grid box perturbation (e.g., 2 km$\times$2 km $\times$0.001 km for the MISMIP+ perturbation experiments) and $k$ is a power-law index that allows for the possiblity of a nonlinear relationship between ice shelf buttressing and the change in GLF (see also \cite{schoof2007}). Here, we use $k=1/n$ with $n=3$.

Despite the existence of many possible factors linking GLF to ice shelf properties (ice flow direction, horizontal gradients in ice shelf geometry, stress fields, strain-rate fields, perturbation locations, etc.), here we mainly examine model stress fields and the distance between perturbations and the GL. This is because, as we show below, these factors correlate closely with the sensitivity of changes in GLF to the imposed ice shelf perturbations. To incorporate the local stress field and its buttressing capacity into our analysis, we also calculate a ``buttressing number'', ($N_b$), analagous to that from \cite{furst2016} (Eqn \ref{butN}),
\begin{equation}
N_b=1-\frac{\sigma_{nn}}{N_0},
\label{butN}
\end{equation}
where $N_0$ is the vertically integrated ocean pressure ($N_0=0.5\left(1-{\rho_i}/{\rho_w}\right)gH$) and $\rho_i$ (910 kg m$^3$) and $\rho_w$ (1028 kg m$^3$) are the density of ice and ocean water, respectively. $\sigma_{nn}$ is the normal stress along a specific horizontal direction, which we discuss further below.

@@

\section{Results and Discussions}

%\subsection{Linear relationship between buttressing ($N_b$) and GL flux responses ($N_r$)}
\subsection{Correlation between the buttressing number and changes in GLF}

A decrease in ice shelf buttressing tends to lead to an increase in GLF \citep[e.g.,][]{gagliardini2010} and intuitively we expect that the GLF should be relatively more sensitive to melt changes that occur in regions of relatively larger buttressing. Here, we aim to better understand and quantitfy the relationship between the buttressing ``strength'' (given by $N_r$) and corresponding changes in GLF (given by $N_r$). Can we quantify the predicted change in GLF for a given buttressing number ($N_b$) and melt perturbation? In Figure \ref{fig2}, we show correlations between $N_b$ and $N_r$ for perturbations applied to the MISMIP+ domain when $N_b$ from Equation \ref{butN} sets $\sigma_{nn}$=$\sigma_{p1}$, with $\sigma_{p1}$ being the first principle stress\footnote{We expand on our reasons for choosing $\sigma_{p1}$ below.}. In Figure \ref{fig2}b, we include all perturbed ice shelf locations in the comparison and find relatively week $N_b$-$N_r$ correlations. In Figure \ref{fig2}c, we remove from the comparison points that are 1) weakly buttressed ($x>480$, where the ice shelf becomes unconfined) and 2) within 5 km of the GL, and a much stronger $N_b$-$N_r$ correlation emerges. In Figure \ref{fig2}d, we plot the $N_b$-$N_r$ correlation for locations where $x>480$ and the magnitude of the thickness gradient is $<7$x$10^{-3}$. The strong similarities between Figures \ref{fig2}c and \ref{fig2}d suggest that the seemingly \textit{ad hoc} ``distance from the grounding line'' constraint is important for removing areas of complex geometry (and hence complex dynamics) from the $N_b$-$N_r$ comparisons.

% in Equation \ref{butN}) in the MISMIP+ test case (Fig \ref{fig2}a), we find relatively weak $N_b$-$N_r$ correlations when we include the results from all perturbation experiments (Fig \ref{fig2}b). However, if we do not consider the perturbation points that are weakly buttressed ($x>480$ km; near the open shelf region) and are close to GL (the minimum distances between GL and perturbation points are greater than 5 km), we can clearly see a strong linear $N_b$-$N_r$ regression (Fig \ref{fig2}c). A similar strong linearity can also be found if we apply a different filtering approach using the ice thickness gradient field (Fig \ref{fig2}a), i.e., we keep the perturbation points that have thickness gradients smaller than 0.0007 (Fig \ref{fig2}d). This nonlinearity feature is likely caused by the intensive shearing and complex ice flow mechanics near GL (a strong transition zone from relatively slow grounded ice flow to fast extensively floating ice flow). We can again see this non-linearity region in our adjoint sensitivity experiments (Fig. \ref{fig9} in section ``Adjoint sensitivity'').

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/fig2.pdf}
    \caption{(a)MISMIP+ steady-state geometry. Color represents the magnitude of the ice thickness gradient and the white line represents the GL. (b) $N_b$-$N_r$ correlation for all perturbation points. (c) $N_b$-$N_r$ correlation for perturbation points satisfying $x>480$ and that are $>$10 km away from GL; (d) $N_b$-$N_r$ correlation for perturbation points satisfying $x>480$ and with a thickness gradient magnitude of $<7$x$10^{-3}$. In (c) and (d), the red (blue) dots in the insets represent the GL (perturbation) grid cells.}
\label{fig2}
\end{figure}

%\subsection{The buttressing directions}
\subsection{Correlation dependence on buttressing direction}

According to Equation \ref{butN}, the buttressing number $N_b$ is computed using the normal stress ($\sigma_{nn}$) along a specified direction. Therefore, the buttressing number at any perturbation point can vary depending on the chosen direction. \cite{furst2016} calculated $N_b$ along two directions, the ice flow direction and the direction corresponding to the second principle stress ($\sigma_{p2}$), and found that the latter -- the direction corresponding to the maximum compressive stress (or the least extensional stress) -- has the maximum impact on the ``passive'' ice shelf regions. In Figure \ref{fig3}, we plot the correlation coefficients ($r^2$) between $N_r$ and $N_b$ using values for $\sigma_{nn}$ that vary continuously by an angle $\Delta \phi$, between 0 and 180 degrees, relative to the direction of $\sigma_{p1}$. We find the largest correlation coefficient ($r^2>0.9$) when $N_b$ is aligned with the $\sigma_{p1}$ direction ($\Delta \phi=0^\circ$) and the smallest correlation coefficient ($r^2<0.4$) when $N_b$ is aligned with the $\sigma_{p2}$ direction ($\Delta \phi=90^\circ$). 
% for different directions with respect to the first principle stress and flow direction, respectively (Fig \ref{fig3}). Differently, we find that $N_b$ along the $\sigma_{p1}$ direction shows the best regression performance, whereas the $\sigma_{p2}$ appears to show the weakest correlations to $N_r$. This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$).

% As a second, somewhat independent check on the strong correlation between $N_r$ and $N_b$ as a function of the direction $\sigma_{p1}$, we also calculate the correlation coefficient for $N_b$ where the value for $\sigma_{nn}$ is calculated as a function of the ice flow direction. The blue line in Figure \ref{fig3}a is analogous to the red line but for the case where $\Delta \phi$ is relative to the ice flow direction. A similar sinusoidal pattern is evident, but with a phase lag of approximately $50^\circ$ relative to the red line. Thus, aligning the blue curve so that its maximum and minimum coincide with those of the red curve requires rotating the direction for calculating $\sigma_{nn}$ by $50^\circ$.

This can be further testified by looking at the angle differences between $\sigma_{p1}$ and flow directions ($\Delta \phi = \phi_{flow}-\phi_{\sigma_{p1}}$). From Figure \ref{fig3}b, we see that for around 50\% of the perturbation points, their flow directions are around 30--50 degree more than the $\sigma_{p1}$ directions, which is consistent with the phase differences in Figure \ref{fig3}a. If we add around 40 degree on top of the flow direction (blue curve), the new direction will be likely aligned closely with the $\sigma_{p2}$ direction, which is consequently pointing to the smallest $r^2$ number for the $\sigma_{p1}$ (red) curve. This indicates that the local maximum buttressing relating to $\sigma_{p2}$ is unnecessarily corresponding to the integrated instantaneous GLF responses. (\textbf{SP: I tried to reword this so that it is a bit more clear, but I failed (commented out my new version of it). I'm going to leave it for the time being and come back to it later. But I also wonder if this discussion is necessary or if it is just too confusing.})
